Superphy: Real-time prediction of \textit{E. coli} phenotype using graph-based genome sequence storage and retrieval with the \textit{spfy} module

### 2017-11-30 paper-webserver.tex
STOP: need to spend some time chopping through the discussion

\section{RESULTS}
TODO: Update with all of Enterobase
Spfy was tested with 59,5323 public \textit{E. coli} genomes, 5,353 genomes from GenBank and 54,181 genomes from Enterobase totaling 596 GB, storing both the entire sequences and results for all included analysis modules.
The resulting database had XYZ nodes and XYZ edges, with XYZ object properties, which worked out to XYZ GB of data stored.

@@

One of the key goals in developing Spfy is to maintain instantaneity, as modern websites have accustomed users to immediate results.
We attempt to use innovations in web development to bring a similar experience to Spfy as a predictive genomics platform for \textit{E. coli}.
The use of Docker containerization, task queues, and a graph database, were all chosen to support \textit{instantaneity}; modern analytics platforms must respond to user-requested analyses of big-data in the same efficiency as old analyses of single files.

@@

AMR testing, VF testing, and Shiga-toxin testing are of particular importance to surveillance efforts as the targets play a key role in a pathogen's lethality to humans.
Current efforts are focused on predictive genomics, where the relevant phenotypic information can be determined through examination of the whole-genome sequence.
WGS methods allow a single laboratory method - the sequencing step - to provide a variety of analysis options, such as AMR, VF, and Shiga-toxin results, and as such can be used to evaluate the spread of outbreaks with better resolution and context than traditional methods \citep{ronholm2016navigating}.

% application: results similar to a wet-lab
Spfy uses WGS results.
After initial sequencing of new isolates, Spfy can be used in place of a traditional reference laboratory, to determine the O-type and H-type, Stx type, and all known VFs and AMR genes in real-time.
These results can be shared with other agencies and researchers over the internet.
Furthermore, using Spfy's database of pre-processed genomes, Spfy can determine all strains a sample may be related to which is useful for monitoring the evolution of pathogens over time.
A computational approach saves the time and cost associated with performing multiple tests per sample, and allows population comparisons which would not be possible with the current pace of WGS results.

@@

% namely galaxy
Other scientific workflow technologies such as Galaxy \citep{goecks2010galaxy}, and bioinformatics pipelines IRIDA and BAP are used to run analysis modules on WGS data.
Galaxy aims to provide a reproducible computation-based research environment which is accessible to individuals without programming knowledge. Galaxy tackles the problem of linking different analysis software together, by defining interdependencies using a custom schema. A visual workflow editor is also provided for ease of use.
IRIDA is build on top of the Galaxy framework, and adds prebuilt pipelines specific to bioinformatics uses, as well as sequence and result storage. IRIDA takes a project-based approach, with sequences stored per project, and results store linearly per sequence. Focus is given on refined controls for collaborating on projects, and common terms found in the GenEpiO ontology are used to describe results.

The Bacterial Analysis Platform (BAP), developed out of the Technical University of Denmark provides an integrated analysis pipeline for bacterial WGS data as a web service.
BAP takes a combined approach to genome analysis including different programs, such as for VF and AMR determination, by default into the pipeline.
A record of the different files generated by an analysis are stored in a relational database, and the record is queried to determine which files are returned to the user \citep{thomsen2016bacterial}.

>>>>

\section{RESULTS}
TODO: Update with all of Enterobase
Spfy was tested with 59,5323 public \textit{E. coli} genomes, 5,353 genomes from GenBank and 54,181 genomes from Enterobase (\~596 GB), storing both the entire sequences and results for all included analysis modules.
The resulting database had XYZ million nodes and XYZ million edges, with XYZ object properties, which worked out to XYZ TB of data stored.

@@

One of the key goals in developing Spfy is to maintain instantaneity, as modern websites have accustomed users to immediate results.
We attempt to use innovations in web development to bring a similar experience to Spfy as a predictive genomics platform for \textit{E. coli}.
The use of Docker containerization, task queues, and a graph database, were all chosen to support instantaneity; modern analytics platforms must respond to user-requested analyses of big-data in the same efficiency as old analyses of single files.

@@

AMR testing, VF testing, and Shiga-toxin testing are of particular importance to surveillance efforts due to their role in a pathogen's lethality.
Current efforts are focused on predictive genomics, where the relevant phenotypic information can be determined through examination of the whole-genome sequence.
WGS methods allow a single laboratory method - the sequencing step - to provide a variety of analysis options, such as AMR, VF, and Shiga-toxin results, and as such can be used to evaluate the spread of outbreaks with better resolution and context than traditional methods \citep{ronholm2016navigating}.

% application: results similar to a wet-lab
Spfy uses WGS results.
After initial sequencing of new isolates, Spfy can be used in place of a traditional reference laboratory, to determine the O-type and H-type, Stx type, and all known VFs and AMR genes in real-time.
These results can be shared with other agencies and researchers over the internet.
Furthermore, using Spfy's database of pre-processed genomes, Spfy can determine all strains a sample may be related to which is useful for monitoring the evolution of pathogens over time.
A computational approach saves the time and cost associated with performing multiple tests per sample, and allows population comparisons to keep pace the current generation WGS data.

@@

% namely galaxy
Other scientific workflow technologies such as Galaxy \citep{goecks2010galaxy}, and bioinformatics pipelines IRIDA and BAP also run analysis modules on WGS data.
Galaxy aims to provide a reproducible, computation-based research environment which is accessible to individuals without programming knowledge. Galaxy tackles the problem of linking different analysis software together, by defining interdependencies using a custom schema. A visual workflow editor is also provided for ease of use.
IRIDA is build on top of the Galaxy framework, and adds prebuilt pipelines specific to bioinformatics uses, as well as sequence and result storage. IRIDA takes a project-based approach, with sequences stored per project, and results stored linearly per sequence. IRIDA adds Controls for collaborating on projects, and uses common terms found in the GenEpiO ontology to describe results.

The Bacterial Analysis Platform (BAP), developed out of the Technical University of Denmark, provides an integrated analysis pipeline for bacterial WGS data as a web service.
BAP takes a combined approach to genome analysis including different programs, such as for VF and AMR determination, by default into the pipeline.
A record of the different files generated by an analysis are stored in a relational database, and the record is queried to determine which files are returned to the user \citep{thomsen2016bacterial}.
